---
title: "Optimization-HW2-FangZheng"
author: "FangZheng"
date: "2018/2/6"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1
###(a)
Since for every $ x_i, i = 1, 2,..., n $, there are
$$
p(x_i, \Theta) = \frac{1}{\pi[1+(x_i-\Theta)]}
$$
So
$$
l(\theta)=\ln(\prod_{i=1}^{n}{p(x_i,\Theta)})=\sum_{i=1}^{n}{\ln(p(x_i,\Theta))}=-n\ln(\pi)-\sum_{i=1}^{n}{\ln(1+(\Theta-x)^2)}
$$
Take derivatives on both sides
$$
l'(\theta)= -2\sum_{i=1}^{n}{\frac{\theta-x_i}{1+(\Theta-x)^2}}
$$
Take second derivatives on both sides
$$
l''(\theta)= -2\sum_{i=1}^{n}{\frac{1-(\theta-x_i)^2}{(1+(\Theta-x)^2)^2}}
$$
We alredy kown
$$
P(x)=\frac{1}{\pi(1+x^2)} ;
P'(x)=-\frac{1}{\pi}~\frac{2x}{(1+x^2)^2} ;
P'^2(x)=\frac{4}{\pi^2}~\frac{x^2}{(1+x^2)^4} 
$$
So
$$
I(\theta)=n\int\frac{{p'(x)}^2}{p(x)}dx=\frac{4n}{\pi}\int_{-\infty}^\infty\frac{x^2dx}{(1+x^2)^3}
$$
Let $x=\tan\beta$, then we have  $1+\tan^2\beta=\arcsin^2\beta=\frac{1}{\cos^2\beta}$ and $dx=\frac{d\beta}{\cos^2beat}$
Then
$$
I(\theta)=\frac{4n}{\pi}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\frac{\tan^2\beta}{(1+tan^2\beta)^3}\frac{1}{\\cos^2\beta}d\beta=\frac{4n}{\pi}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^2\beta\cos^2\beta~d\beta=\frac{4n}{\pi}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^2\beta(1-\sin^2\beta)~d\beta=\frac{4n}{\pi}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^2\beta-\sin^4\beta~d\beta
$$

As $1-\cos2\beta=2\sin^2\beta$, then we have

$$
\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^2\beta~d\beta=\frac{1}{2}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}(1-\cos2\beta)d\beta=\frac{1}{2}(\beta-\frac{1}{2}\sin2\beta)|_{-\frac{\pi}{2}}^{\frac{\pi}{2}}~=\frac{1}{2}
$$

$$
\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\sin^4\beta~d\beta=\frac{1}{4}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}(1-\cos2\beta)^2d\beta=\frac{1}{4}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\cos^22\beta-2\cos2\beta+1~d\beta=
$$
$$
=\frac{1}{4}\int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}\frac{\cos4\beta}{2}-2\cos2\beta+\frac{3}{2}~d\beta=\frac{1}{4}(\frac{1}{8}\sin4\beta-\sin2\beta+\frac{3}{2}\beta)|_{-\frac{\pi}{2}}^{\frac{\pi}{2}}~=\frac{3\pi}{8}
$$

<font size = 3>*Finally we have:* </font>
$$
I(\theta)=\frac{4n}{\pi}(\frac{\pi}{2}-\frac{3\pi}{8})=\frac{n}{2}
$$



###(b)
```{R}
x1=c(1.77,-0.23,2.76,3.80,3.47,56.75,-1.34,4.24,-2.44,3.29,3.71,-2.40,4.53,-0.07,-1.05,-13.87,-2.53,-1.75)

n = length(x1)

l <- function(theta) {
  return(n*log(pi) + sum(log(1+(theta-x1)^2))) #-log
}

l.grid <- function(theta) {
  return(2*sum((theta-x1)/(1+(theta-x1)^2))) #-l.grid
}

l.hess <- function(theta) {
  return(matrix(2*sum((1-(theta-x1)^2)/(1+(theta-x1)^2)^2),nrow=1)) #-l.hess
}

theta.est <- function(starting) {
  ## MLE  
  theta <- nlminb(start=starting,l,l.grid,l.hess)$par
  return(theta)
}



# plot

theta.value = seq(-11, 38, by=0.05)

log.likelihood = c()

for (i in 1:length(theta.value)){
  log.likelihood[i] = -l(theta.value[i])
}

plot(theta.value, log.likelihood, xlab="parameter.theta",ylab="log.likelihood", type="l")

theta.est(-11)
theta.est(-1)
theta.est(0)
theta.est(1.5)
theta.est(4)
theta.est(4.7)
theta.est(7)
theta.est(8)
theta.est(38)
```



###(c)
```{R}
theta.est.fix <- function(starting) {
  alpha <- c(1,0.64,0.25)
  ## MLE
  theta <- c()
  for (j in 1:3){
    l.hess.fix <- function(theta) {
      return(matrix(1/alpha[j],nrow=1))
    }
    theta[j] <- nlminb(start=starting,l,l.grid,l.hess.fix)$par
  }
  return(theta)
}
theta.est.fix(-11)
theta.est.fix(-1)
theta.est.fix(0)
theta.est.fix(1.5)
theta.est.fix(4)
theta.est.fix(4.7)
theta.est.fix(7)
theta.est.fix(8)
theta.est.fix(38)
```



###(d)
```{r}
theta.est.fisher <- function(starting){
  ## MLE  
  theta <- starting
  delta <- 1
  while(abs(delta) >= 0.00001){
    theta1 <- theta 
    dl <- l.grid(theta)
    theta <- theta - dl/(n/2)
    delta <- theta - theta1
  }
  return(theta)
}

theta.est.fisher(-11)
theta.est.fisher(-1)
theta.est.fisher(0)
theta.est.fisher(1.5)
theta.est.fisher(4)
theta.est.fisher(4.7)
theta.est.fisher(7)
theta.est.fisher(8)
theta.est.fisher(38)
```



###(e)
Based on the results above, blow is the table that shows the rank of some feature of different method.
```{R}
knitr::kable(
  matrix (c (1, 3, 2, 1, 3, 1, 3, 1, 2), nrow = 3, ncol = 3, byrow = TRUE, dimnames = list(c("speed",  "stability", "accuracy"), c("method_b", "method_c", "method_d"))) 
)
```



## Q2
###(a)
```{R}
x2 = c(3.91,4.85,2.28,4.06,3.70,4.04,5.46,3.53,2.28,1.96,2.53,3.88,2.22,3.47,4.82,2.46,2.99,2.54,0.52)
a.value = seq(-pi, pi, by=0.05)
n = length(x2)
log.like = c()
for (i in 1:length(a.value)){
  log.like[i] = -n*log(2*pi)+sum(log(1-cos(x2-a.value[i])))
}
plot(a.value, log.like, xlab="parameter.theta",
     ylab="log.likelihood", type="l")
```



###(b)
```{R}
a.est <- function(sample,starting) {
  n = length(sample)
  
  l <- function(a) {
    return(-n*log(2*pi) + sum(log(1-cos(sample-a)))) #log
  }
  
  l.grid <- function(a) {
    return(sum(sin(a-sample)/(1-cos(a-sample)))) #l.grid
  }
  
  l.hess <- function(a) {
    return(sum(1/(cos(a-sample)-1))) #l.hess
  }
  
  ## MLE  
  theta <- starting
  delta <- 1
  while(abs(delta) >= 1e-10){
    theta1 <- theta 
    dl <- l.grid(theta)
    ddl <- l.hess(theta)
    if (ddl==0){
      return(NULL)
    }
    theta <- theta - dl/ddl
    delta <- theta - theta1
  }
  
  return(c(theta,l(theta)))
}

m = mean(x2)
theta0 = asin(m-pi)
theta0
```



###(c)
```{r}
a.est1 = a.est(x2,theta0)
a.est1[1]
```



###(d)
```{r}
a.est2 = a.est(x2,-2.7)
a.est3 = a.est(x2,2.7)
a.est2[1]
a.est3[1]
```



###(e)
```{r}
a.value = seq(-pi,pi,by = 2*pi/199); # set 200 equally spaced starting point
a.par = c()
a.mle = c()
for (i in 1:length(a.value)){
  a.par[i] = a.est(x2,a.value[i])[1]
  a.mle[i] = a.est(x2,a.value[i])[2]
}

ascend = order(a.mle)
a.mle.ascend = a.mle[ascend]
a.par.ascend = a.par[ascend]

# find the unique outcome
j = 1
p = 2
i = 1
e = c()
e[1] = a.mle.ascend[1]
while(j<=200){
  if(a.mle.ascend[i]!=a.mle.ascend[j]){
    e[p] = a.mle.ascend[j]
    p = p+1
    i = j
    j = j+1
  }else{j = j+1}
}

# classify the starting point according to the unique outcome
group = matrix(nrow = length(e),ncol = 46)
for (y in 1:length(e)){
  t = 1
  for (i in 1:200){
    if(a.mle.ascend[i]== e[y]){
      group[y,t] = round(a.value[ascend[i]],5)
      t = t+1
    }
  }
}
group[is.na(group)]=""
knitr::kable(group)
```



##Q3
###(a)
```{r}
t <- c(0,8,28,41,63,69,97,117,135,154)
x <- c(2,47,192,256,768,896,1120,896,1184,1024)
f <- expression(2*K/(2+(K-2)*exp(-r*t)))

df <- function(K,r,t){
  dfk <- D(f,"K")
  dfr <- D(f,"r")
  K <- K 
  r <- r 
  t <- t
  a <- eval(dfk)
  b <- eval(dfr)
  c <- array(c(a,b),c(1,2))
  return(c)
}

Df <- function(K,r){
  a <- K
  b <- r
  m <- df(a,b,t[1])
  for(i in 2:10){
    c <- df(a,b,t[i])
    m <- rbind(m,c)
  }
  return(m)
}

Z <- function(K,r){
  a <- c()
  for(i in 1:10){
    a[i] <- x[i] - 2*K/(2+(K-2)*exp(-r*t[i]))
  }
  m <- array(a,c(10,1))
  return(m)
}

theta <- matrix(c(1200,0.2),nrow=2)
delta <- matrix(c(1,1),nrow=2)

while(crossprod(delta,delta)>=0.001){
  theta1 <- theta 
  a <- Df(theta[1,1],theta[2,1])
  z <- Z(theta[1,1],theta[2,1])
  theta <- theta + solve(t(a)%*%a)%*%t(a)%*%z
  delta <- theta - theta1
}

a.est <- theta
print(a.est)
```



#(b)
```{r}
f <- function(K,r){
  return(sum((x-2*K/(2+(K-2)*exp(-r*t)))^2))
}

z <- matrix(0,100,100,byrow=T)
for (i in 1:100){
  for (j in 1:100){
    K <- 600 + 8*j
    r <- 0 + 0.01*i
    z[j,i] <- f(K,r)
  }
}
K <- seq(600,1400,length.out = 100)
r <- seq(0,1,length.out = 100)
contour(K,r,z,xlab="K",ylab="r")
```



#(c)
```{r}
l <- expression(log(1/(sqrt(2*pi)*sigema))-
                  (log((2*2+2*(K-2)*exp(-r*0))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*47+47*(K-2)*exp(-r*8))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*192+192*(K-2)*exp(-r*28))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*256+256*(K-2)*exp(-r*41))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*768+768*(K-2)*exp(-r*63))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*896+896*(K-2)*exp(-r*69))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*1120+1120*(K-2)*exp(-r*97))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*896+896*(K-2)*exp(-r*117))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*1185+1184*(K-2)*exp(-r*135))/(2*K)))^2/(2*sigema^2)+
                  log(1/(sqrt(2*pi)*sigema))-
                  (log((2*1024+1024*(K-2)*exp(-r*154))/(2*K)))^2/(2*sigema^2))

dl <- function(beita){
  dlk <- D(l,"K")
  dlr <- D(l,"r")
  dlsigema <- D(l,"sigema")
  K <- beita[1] 
  r <- beita[2] 
  sigema <- beita[3]
  a <- eval(dlk)
  b <- eval(dlr)
  c <- eval(dlsigema)
  return(c(a,b,c))
}

ddl <- function(beita){
  dlkk <- D(D(l,"K"),"K")
  dlkr <- D(D(l,"K"),"r")
  dlksigema <- D(D(l,"K"),"sigema")
  dlrr <- D(D(l,"r"),"r")
  dlrsigema <- D(D(l,"r"),"sigema")
  dlsigema2 <- D(D(l,"sigema"),"sigema")
  K <- beita[1] 
  r <- beita[2] 
  sigema <- beita[3]
  a <- c(eval(dlkk),eval(dlkr),eval(dlksigema),eval(dlkr),eval(dlrr),
         eval(dlrsigema),eval(dlksigema),eval(dlrsigema),eval(dlsigema2))
  m <- matrix(a,byrow=TRUE,nrow=3)
  return(m)
}

a <- matrix(c(1200,0.2,0.5),nrow=3)
delta <- matrix(c(1,1,1),nrow=3)
while(crossprod(delta,delta)>=0.001){
  b <- a
  c <- matrix(dl(a),nrow=3)
  d <- solve(ddl(a))
  a <- a - d%*%c
  delta <- a - b
}
a
solve(-ddl(a))
```